---
title: "TEAM"
output: html_document
---

# Loading the Data Sets in to Rstudio
```{r}
library(readxl)
library(tidyverse)
library(dplyr)
Defense <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2019.xlsx', sheet='Defensive')
str(Defense)
Offense <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2019.xlsx', sheet='Offensive')
str(Offense)
Biography <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2019.xlsx', sheet='Biography')
str(Biography)
```

# Cleaning for Defense
```{r}
defClean <- Defense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Tackles_Solo = as.numeric(Tackles_Solo), 
         Tackles_ASST = as.numeric(Tackles_ASST), 
         Tackles_TFL = as.numeric(Tackles_TFL), 
         Tackles_Sack = as.numeric(Tackles_Sack), 
         Turnover_FF = as.numeric(Turnover_FF), 
         Turnover_FR = as.numeric(Turnover_FR), 
         Pass_QBH = as.numeric(Pass_QBH), 
         Pass_PB = as.numeric(Pass_PB), 
         Turnover_INT = as.numeric(Turnover_INT))
str(defClean)
```

# Cleaning for Offense
```{r}
Offense <- Offense %>%
  separate(Passing_CMPATT, into = c("Passing_CMP", "Passing_ATT"))
str(Offense)
  
  
offClean <- Offense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT),
         Passing_CMP = as.numeric(Passing_CMP),
         Passing_ATT = as.numeric(Passing_ATT))
str(offClean)
```


# Cleaning for Biography
```{r}
Biography <- Biography %>%
  separate(Height, into = c("Height_Feet", "Height_Inches"))

bioClean <- Biography %>%
  mutate(Height_Feet = as.numeric(Height_Feet),
         Height_Inches = as.numeric(Height_Inches))
str(bioClean)
```


# Tidying - 1
```{r}
defClean <- defClean %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to = 'DEFstat', values_to = 'DEFstat_count')
head(defClean)

offClean <- offClean %>%
  pivot_longer(Rushing_ATT:Passing_INT, names_to = 'OFFstat', values_to = 'OFFstat_count')
(offClean)
```

# Tidying - 2
```{r}
library(ggplot2)
ggplot(defClean, aes(x = DEFstat, weight = DEFstat_count)) + geom_bar() + coord_flip()
```


# Tidying -3
```{r}
Iowa <- defClean %>%
  select(Opponent_Opponent, DEFstat, DEFstat_count) %>%
  filter(Opponent_Opponent %in% c('Iowa'), DEFstat %in% 'Tackles_Solo') 
head(Iowa)
NotreDame <- defClean %>%
  select(Opponent_Opponent, DEFstat, DEFstat_count) %>%
  filter(Opponent_Opponent %in% c('Notre Dame'), DEFstat %in% 'Tackles_Solo') 
head(NotreDame)

Opponent <- Iowa %>% full_join(NotreDame)
Opponent <- Opponent %>%
  select(Opponent_Opponent, DEFstat_count)
head(Opponent)
ggplot(Opponent, aes(x = Opponent_Opponent, y = DEFstat_count)) + geom_point() + geom_jitter()  
```

## ISU had a better defensive game against Notre Dame because they had more players who had more solo tackles.



# Tidying - 4
```{r}
bioClean <- bioClean %>%
  separate(Hometown, into = c("City", "State"), sep= ',')
head(bioClean)
```
# Tidying - 5
```{r}
q5 <- bioClean %>%
  group_by(State) %>%
summarize(Total = n())
(q5)
```


# Joining - 1
```{r}
TacklesdefClean <- defClean %>%
  filter(DEFstat == "Tackles_Solo") %>%
  group_by(Name) %>%
  summarize(total_tackles = sum(DEFstat_count)) %>%
  arrange(total_tackles)
TacklesdefClean

OffBio <- offClean %>% left_join(bioClean) 
OffBio

total2019 <- OffBio %>% left_join(defClean)
view(total2019)

offstates <- OffBio %>%
  group_by(State) %>%
summarize(Total = n())

offstates

```

# Joining - 2
```{r}

```



# Joining - 3
## Adding 2018 data
```{r}
OldDefense <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2018.xlsx', sheet='Defensive')
OldOffense <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2018.xlsx', sheet='Offensive')
OldBiography <- read_excel('Desktop/DS202_Lab4/cyclonesFootball2018.xlsx', sheet='Biography')
```


## Cleaning Defensive 2018 Data
```{r}
OlddefClean <- OldDefense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Tackles_Solo = as.numeric(Tackles_Solo), 
         Tackles_ASST = as.numeric(Tackles_ASST), 
         Tackles_TFL = as.numeric(Tackles_TFL), 
         Tackles_Sack = as.numeric(Tackles_Sack), 
         Turnover_FF = as.numeric(Turnover_FF), 
         Turnover_FR = as.numeric(Turnover_FR), 
         Pass_QBH = as.numeric(Pass_QBH), 
         Pass_PB = as.numeric(Pass_PB), 
         Turnover_INT = as.numeric(Turnover_INT))
str(OlddefClean)

OlddefClean <- OlddefClean %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to = 'DEFstat', values_to = 'DEFstat_count')
head(OlddefClean)

```



## Cleaning Offensive 2018 Data
```{r}
OldOffense <- OldOffense %>%
  separate(Passing_CMPATT, into = c("Passing_CMP", "Passing_ATT"))
str(Offense)
  
  
OldoffClean <- OldOffense %>%
  mutate(Name = as.factor(Name),
         Opponent_Opponent = as.factor(Opponent_Opponent),
         Rushing_ATT = as.numeric(Rushing_ATT),
         Rushing_YDS = as.numeric(Rushing_YDS),
         Rushing_TD = as.numeric(Rushing_TD),
         Receiving_REC = as.numeric(Receiving_REC),
         Receiving_YDS = as.numeric(Receiving_YDS),
         Receiving_TD = as.numeric(Receiving_TD),
         Passing_YDS = as.numeric(Passing_YDS),
         Passing_TD = as.numeric(Passing_TD),
         Passing_INT = as.numeric(Passing_INT),
         Passing_CMP = as.numeric(Passing_CMP),
         Passing_ATT = as.numeric(Passing_ATT))
str(OldoffClean)

OldoffClean <- OldoffClean %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to = 'OFFstat', values_to = 'OFFstat_count')
head(OldoffClean)

```


## Cleaning Biography 2018 Data
```{r}
OldBiography

OldBiography <- OldBiography %>%
  separate(Height, into = c("Height_Feet", "Height_Inches"))

OldbioClean <- OldBiography %>%
  mutate(Height_Feet = as.numeric(Height_Feet),
         Height_Inches = as.numeric(Height_Inches))

OldbioClean <- OldbioClean %>%
  separate(Hometown, into = c("City", "State"), sep= ',')
head(OldbioClean)

```
```{r}
OffBio2018 <- OldbioClean %>% left_join(OldoffClean)
OffBio2018
total2018 <- OffBio2018 %>% left_join(OlddefClean)
total2018

newtotal2018 <- total2018 %>% 
  select(Name, OFFstat_count) %>%
  group_by(Name) %>%
  summarize(total = sum(OFFstat_count))
newtotal2018
  
newtotal2019 <- total2019 %>%
  select(Name, OFFstat_count) %>%
  group_by(Name) %>%
  summarize(total = sum(OFFstat_count))
newtotal2019


```